\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{makeidx}
\usepackage{graphicx}
\author{Group 9}
\begin{document}
	\section{Times Series Analysis}
	
	\section{Classification Methods Use in Other Countries}
	LULC data are records that documents to what extent a region is covered by wetlands, forests, agriculture, impervious surfaces, and 	other land and water forms. These water forms include open water  or wetlands. Land use shows how people use landscape either for conservation, development, agriculture or mixed uses [6, 7]. Changes In land can be identified by analysing satellite imagery. However, land use cannot be identified from satellite imagery. Satellite imagery give us  information that helps in understanding the present landscape. Furthermore, to see changes	through time, different years are needed. With this information, we can assess  decades of data as well as  insight into the possible effects of these changes that has occured and make better decisions before they can cause great harm.	According to [10], five defferent types of LULC pattern were classified barren lands such as Galamsey Site, agricultural lands, urban lands, quarries, and free water bodies, to detect the 25years LULC change in the western Nile delta of Egypt. Supervised maximum likelihood classification (MLC)	method together with landsat images were used in Erdas Imagine software. The finding shows a significant
	change in barren land changing into agricultural land continuously	from 1984 to 2009.\\
	
	 [10] also used the Maximum likelihood classification(MLC), for  Change detection and spatial matrix analysis to find  changes in Avellino 	Italy for the period (1954 to 2004). The result was a LULC classification, with urban land use rapidly increasing and affecting the cultivated land, while woodland and grassland cover decrease was at a lower rate.\\
	
	
	Moreover, in [13] studied the land use, land cover change in Duzce plain Turkey. They use Supervised classification and the Corine land cover nomenclature methods. The result shows  classes of urban	fabric, forest, heterogeneous agricultural land, inland wasteland	and (Industrial, commercial, and transport) units with an accuracy	assessment between 92.41 $\%$ and 97.3 $\%$ for 2010 and
	1987 respectively. Also, a significant change in LULC was noticed with 11.2$\%$ increase in agricultural area and 335$\%$ decrease	of forest land.\\
	

	
	Furthermore, in [17] study the LULC change of watershed in Pakistan from 1992 to 2012 using the supervised classification of	maximum likelihood algorithm in Erdas Imagine. The finding	shows agriculture, bare soil/rocks, settlements,	vegetation and water. Also, the water body and vegetation are	decreasing in favour of settlements, agriculture and bare soil rapidly from 38.2% and 74.3% respectively.\\
	
	Also, in [18] study, both unsupervised  and supervised  methods ISODATA ana MLA respectivelly  were used for classification. Change	detection and Markov change analysis methods were aslo used to measure the changes generated  in 	Mansoura and Talkha of Egypt from 1985 to 2010. The finding shows classes this is to say agriculture, barren land, built-up	area and water body. Also, a significant change was noticed in	agricultural land and built-up area to tune of 33$\%$ decrease and	30$\%$ increase respectively, while barren land and water bodies
	changes were minimal.\\
	
	Similarly, in [19] studied the LULC classification of Sawantwadi taluka, in India. The hybrid, parametric (MLA and ISODATA),	and nonparametric (DT) methods were used. The finding shows	the classified LULC of the forest, water, built-up, agriculture, 	plantation, fallow land, open and dense shrubland, stone quarry,and grassland with an accuracy assessment of 93$\%$ and koppa of	0.92.\\
	
	Also, in [20] they  measured the land use ,land cover changse in Seramban. In the	study, Natural Breaks (Jenks) and Normalized Difference Vegetation Index (NDVI) methods were used for classification and difference from 1990 to 2000. The result shows four classes of	LULC viz barren land, built-up area, vegetation and water body. A 13$\%$ decrease in vegetation cover was noticed while other land use land cover increase by 3.7$\%$ accordingly with an accuracy	assessment of 87$\%$ and 88$\%$ respectively.\\
	

	
\end{document}